---
title: "Talent Management"
subtitle: "Looking at data on 870 different employees for attrition and salary trends/predictions."
author: "Kendall Scott"
date: "4/16/2022"
output:
  html_document:
    theme: united
    highlight: tango
---
```{r}
library(ggplot2)
library(dplyr)
library(viridis)
library(stringr)

casestudy2 = read.csv(".//CaseStudy2-data.csv") #casestudy2-data.csv

#Attrition by Department
ggplot(casestudy2, aes(x=as.factor(Department), fill=Attrition))+
  geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..]), position="stack" , width=0.5) +
  geom_text(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], label=scales::percent(..count../tapply(..count.., ..x.. ,sum)[..x..]) ),
            stat="count", position=position_stack(0.9), vjust=0.5)+
  xlab('Department') +
  ylab('Percent of Attrition')+
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10))+
  theme(axis.text = element_text(size = 7))

casestudy2$Attritioncalc=case_when(
  casestudy2$Attrition =='Yes'  ~ 1,
   TRUE ~ 0
)

#summary
er<-casestudy2 %>% group_by(Department) %>% summarize(meanincome = mean(MonthlyIncome), calcAttrition = (sum(Attritioncalc)/n()), Employees = n()) %>% arrange(desc(Employees))
er 

#Attrition % by Job Role and Department
#summary table
er<-casestudy2 %>% group_by(Department, JobRole) %>% summarize(meanincome = mean(MonthlyIncome), calcAttrition = (sum(Attritioncalc)/n()), Employees = n()) %>% arrange(desc(Employees))
er 
#graph
ggplot() +
  geom_polygon(data =  er, aes(x=Department, y = JobRole),color = "white", fill="grey", alpha=0.5) +
  geom_point( data=er, aes(x=Department, y=JobRole, color=Employees, size=calcAttrition, alpha=0.5)) +
  scale_color_viridis(option="viridis",  name="Employees" ) +
  scale_size(range = c(1, 10), name="Attrition %")+
  ggtitle("Attrition by Role")

#Attrition % by Job Level and Department
#summary
er<-casestudy2 %>% group_by(Department, JobLevel) %>% summarize(meanincome = mean(MonthlyIncome), calcAttrition = (sum(Attritioncalc)/n()), Employees = n()) %>% arrange(desc(Employees))
er 

#graph
ggplot() +
  geom_point( data=er, aes(x=Department, y=JobLevel, color=Employees, size=calcAttrition,  alpha=0.5)) +
  scale_color_viridis(option="viridis",  name="Employees" ) +
  scale_size(range = c(1, 10), name="Attrition %")+
  ggtitle("Attrition by Job Level")
#summary
er<-casestudy2 %>% group_by(Department, JobLevel) %>% summarize(meanincome = mean(MonthlyIncome), calcAttrition = (sum(Attritioncalc)/n()), Employees = n()) %>% arrange(desc(Employees))
er 

#graph
ggplot() +
  geom_point( data=er, aes(x=Department, y=JobLevel, color=Employees, size=calcAttrition,  alpha=0.5)) +
  scale_color_viridis(option="viridis",  name="Employees" ) +
  scale_size(range = c(1, 10), name="Attrition %")+
  ggtitle("Attrition by Job Level")
```
### By Department
##### The Sales Department has the highest % of Attrition (21.6%)
##### 54% of Attrition comes from Research and Development, but Research and Development is the largest Department.

### By Job Level
##### Overall Job Level 1 has the highest % of Attrition
##### Within the Sales Department– JobLevel 5 has higher attrition % (29%) than Job Level 1 in other departments.

### By Job Role

#### 68% of attrition is from 3 Job Roles:
#### -Sales Executive (Sales)
#### -Research Scientist (Research & Development)
#### -Laboratory Technician (Research & Development)




```{r}
# k-NN or naive Bayes but may also use other models (logistic regression, random forest, LDA, SVM, etc) 
#as long as you compare the results between the two or more models. 
#You may then use any of the models to fulfill the 60/60 sensitivity/specificity requirement. 
#This goes for regression as well; you must use linear regression but may include additional models for comparison and use in the competition (LASSO, random forest, ensemble models, etc.).
casestudy2.noattrition = read.csv(".//CaseStudy2CompSet No Salary.csv")

library(class)
library(caret)
library(e1071)

clean_casestudy2 = data.frame(   
  Attrition=casestudy2$Attrition,
  Age = scale(casestudy2$Age), 
  JobInvolvement=scale(casestudy2$JobInvolvement),
  JobLevel=scale(casestudy2$JobLevel),
  Distance=scale(casestudy2$DistanceFromHome),
  StockOptionLevel=scale(casestudy2$StockOptionLevel),
  EnvironmentSatisfaction=scale(casestudy2$EnvironmentSatisfaction),
  RelationshipSatisfaction=scale(casestudy2$RelationshipSatisfaction),
  JobSatisfaction=scale(casestudy2$JobSatisfaction),
  YearSinceLastPromotion=scale(casestudy2$YearsSinceLastPromotion),
  YearsinCurrentRole=scale(casestudy2$YearsInCurrentRole),
  Education=scale(casestudy2$Education)
)
casestudy2.noatt= data.frame(    
  Attrition=casestudy2.noattrition$Attrition,
  Age = scale(casestudy2.noattrition$Age), 
  JobInvolvement=scale(casestudy2.noattrition$JobInvolvement),
  JobLevel=scale(casestudy2.noattrition$JobLevel),
  Distance=scale(casestudy2.noattrition$DistanceFromHome),
  StockOptionLevel=scale(casestudy2.noattrition$StockOptionLevel),
  EnvironmentSatisfaction=scale(casestudy2.noattrition$EnvironmentSatisfaction),
  RelationshipSatisfaction=scale(casestudy2.noattrition$RelationshipSatisfaction),
  JobSatisfaction=scale(casestudy2.noattrition$JobSatisfaction),
  YearSinceLastPromotion=scale(casestudy2.noattrition$YearsSinceLastPromotion),
  YearsinCurrentRole=scale(casestudy2.noattrition$YearsInCurrentRole),
  Education=scale(casestudy2.noattrition$Education))
      
test= casestudy2.noatt
train=clean_casestudy2

#confusion matrix results for each row
numks = 30

masterAcc = matrix(nrow = numks)
masterSens = matrix(nrow = numks)
masterSpec = matrix(nrow = numks)
masterK = matrix(nrow = numks)

data(attrition)
set.seed(1)
i=1
for(i in 1:numks)
{
 classifications = knn(train[,c(2:12)],test[,c(2:12)],as.factor(train$Attrition), prob = TRUE, k = i)
  #results for accuracty, sensitivity, and specificity
  table(as.factor(test$Attrition),classifications)
  CM = confusionMatrix(table(as.factor(test$Attrition),classifications))
  masterAcc[i] = CM$overall[1]
  masterSens[i]=CM[["byClass"]][["Sensitivity"]][1]
  masterSpec[i]=CM[["byClass"]][["Specificity"]][1]
  masterK[i]=i
}
Overall=cbind(masterAcc,masterSens, masterSpec, masterK)
Overall=as.data.frame(Overall)
#renaming CM column results
Overall=rename(Overall, Accuracy=V1, Sensitivity=V2, Specificity=V3,K=V4)

MeanAcc = colMeans(Overall)
MeanAcc

library("tidyverse")
df <- Overall %>%
  select(K, Accuracy, Sensitivity, Specificity) %>%
  gather(key = "variable", value = "value", -K)

#graph of knn results
ggplot(df, aes(x = K, y = value)) + 
  geom_line(aes(color = variable, linetype = variable)) + 
  scale_color_manual(values = c("black", "steelblue", "blue"))+ geom_vline(xintercept=16,color="darkgreen")+ ylab('Confusion Matrix Measurments')+ggtitle("KNN Results")
#based on results-- k=16 is the best selection

casestudy2.noattrition = read.csv(".//CaseStudy2CompSet No Attrition.csv")
casestudy2.noattrition$Attrition= classifications = knn(train[,c(2:12)],test[,c(2:12)],as.factor(train$Attrition), prob = TRUE, k = 16)


```
### Attrition
#### Looking at the 16 closest employees (k=16) produces best results
#### Model accuracy 83.3%
#### Specificity 83.7%
#### Sensitivity  60.0%


### Factors for the model were:
#### -Age
#### -Job Involvement
#### -Job Level
#### -Distance
#### -Stock Option Level
#### -Environment Satisfaction
#### -Relationship Satisfaction
#### -Job Satisfaction
#### -Year Since Last Promotion
#### -Years in Current Role
#### -Education

```{r echo=FALSE}
#Monthly Income
#linear regression

fit =lm(formula = MonthlyIncome~  JobLevel+YearsInCurrentRole, casestudy2)
#fit a regression model and use LOOCV to evaluate performance

summary(fit)
confint(fit)
preds = predict(fit)
RMSE=sqrt(mean(fit$residuals^2))
casestudy2$preds = predict(fit)

er<-casestudy2 %>% group_by(JobLevel, YearsInCurrentRole) %>% summarize(MeanIncome = mean(MonthlyIncome), meanPred = mean(preds))

#Pred vs. Mean Income
er %>% ggplot(aes(x = YearsInCurrentRole, y = MeanIncome)) + geom_point() +geom_line(data = er, aes( x = YearsInCurrentRole, y = meanPred, col = "red"))+facet_grid(rows = vars(JobLevel))

casestudy2.nosal= read.csv(".//CaseStudy2CompSet No Salary.csv")
casestudy2.nosal$MonthlyPred=predict(fit, newdata = casestudy2.nosal)
hist(fit$residuals, col = "blue", main = "Histogram of Residuals")

#mlr and 3d plotting
library(plotly)
library(reshape2)

#plot the regression plane
#Graph Resolution 
graph_reso <- .5

#Setup Axis
axis_x <- seq(min(casestudy2$JobLevel), max(casestudy2$JobLevel), length = 100)
axis_y <- seq(min(casestudy2$YearsInCurrentRole), max(casestudy2$YearsInCurrentRole), length = 100)

#Sample points
surface_inc <- expand.grid(JobLevel = axis_y,YearsInCurrentRole = axis_x,KEEP.OUT.ATTRS = F)
surface_inc$Income <- predict.lm(fit, newdata = surface_inc)
surface_inc <- acast(surface_inc, JobLevel ~ YearsInCurrentRole, value.var = "Income") #y ~ x

surface_plot <- plot_ly(casestudy2, 
                        x = ~JobLevel, 
                        y = ~YearsInCurrentRole, 
                        z = ~MonthlyIncome,
                        text = "Income", 
                        type = "scatter3d",
                        mode = "contours")


surface_plot <- add_trace(p = surface_plot,
                          z = surface_inc,
                          x = axis_x,
                          y = axis_y,
                          type = "surface",
                          mode = "contours")

surface_plot

#Random Forest
library(randomForest)
library(varImp)

clean_casestudy2 = data.frame(   
  MonthlyIncome=casestudy2$MonthlyIncome,
  Gender=as.factor(casestudy2$Gender),
  TrainingTimesLastYear=scale(casestudy2$TrainingTimesLastYear),
  Attrition = as.factor(casestudy2$Attrition),
  Department=as.factor(casestudy2$Department),
  Age = scale(casestudy2$Age), 
  Marriage=as.factor(casestudy2$MaritalStatus),
  JobRole = as.factor(casestudy2$JobRole), 
  JobInvolvement=scale(casestudy2$JobInvolvement),
  JobLevel=scale(casestudy2$JobLevel),
  Distance=scale(casestudy2$DistanceFromHome),
  BusinessTravel = as.factor(casestudy2$BusinessTravel), 
  StockOptionLevel=scale(casestudy2$StockOptionLevel),
  EnvironmentSatisfaction=scale(casestudy2$EnvironmentSatisfaction),
  RelationshipSatisfaction=scale(casestudy2$RelationshipSatisfaction),
  JobSatisfaction=scale(casestudy2$JobSatisfaction),
  OverTime=as.factor(casestudy2$OverTime),
  YearSinceLastPromotion=scale(casestudy2$YearsSinceLastPromotion),
  YearsinCurrentRole=scale(casestudy2$YearsInCurrentRole),
  Education=scale(casestudy2$Education),
  EducationField=as.factor(casestudy2$EducationField),
  TotalWorkingYears=scale(casestudy2$TotalWorkingYears)
)
model <- randomForest(MonthlyIncome ~  ., data = clean_casestudy2, importance=TRUE) 
importance=importance(model)
varImpPlot(model)


```

### Monthly Income
### Used linear regression to predict monthly incoming using:
#### -Job Level
#### -Years in Current Role

### RMSE(root mean squared error) of $1,413

### Attrition Factors
#### Overall the most important factors for Attrition leveraging Random Forest:
#### -Job Level
#### -Job Role
#### -Years in Current Role

